---
title: Manuscript figures
date: Last edited:`r Sys.Date()`
output: 
  word_document: 
    reference_docx: "template_adjusted.docx"
    toc: true
    number_sections: false
    
---

```{r packages, warning=FALSE, message=FALSE, include=FALSE}
# Packages
library(here)
library(tidyverse)
library(ggplot2)
library(ggridges) # for forklength joyplot
library(pals) # check color contrasts
library(patchwork) # align plots and inset figures
library(brms) # import model
library(bayesplot) # for model fit and ppc_check
library(data.table) # fread() for covariate data
library(tidybayes) # gathering draws and epred
library(ggdist) # plotting stat_half_eye color ramp
library(flextable) # use for all table design word or pdf- https://ardata-fr.github.io/flextable-book/index.html#create-a-flextable
library(ftExtra) # use to split headers better then separate_headers() https://cran.r-project.org/web/packages/ftExtra/vignettes/transform-headers.html
library(modelr) # seq_range
```

```{r setup-chunk, echo=FALSE}
knitr::opts_chunk$set(
  fig_cap = TRUE,
  dpi = 300,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.path = here("results/figures/manuscript/")
)
```


```{r color_palette, include=FALSE}
# https://jfly.uni-koeln.de/color/

pal_all <- c("#000000", "#d1ae41", "#56B4E9", "#D55E00", "#0072B2", "#CC79A7", "#009E73", "#F0D449", "#E69F00") 

pal_age <- c("#444444", "#83dcdd", "#e04555", "#000000")

pal_cov <- c("#d1ae41", "#009E73", "#2b92b4") # gray: #555d66

# check contrasts
# pal.safe(pal_cov)
```

```{r func_figbreaks,include=FALSE}
# set function for SYage1 shown on graphs
every_nth <- function(n) {
  return(function(x) {
    x[c(TRUE, rep(FALSE, n - 1))]
  })
}
```

```{r import_data,cache=TRUE, include=FALSE}
# import data
DATA <- read_csv(here("data", "adult_age_covariate_data.csv"))

# scale data
data.scaled <- DATA %>%
  mutate_at(c(16:21), ~ (scale(.) %>% as.vector())) %>%
  # add each OA for figure 2 & 3
  mutate(OA.true = case_when(
    adult_age == 3 ~ "1-Ocean",
    adult_age == 4 ~ "2-Ocean",
    adult_age == 5 ~ "3-Ocean",
    adult_age == 6 ~ "4-Ocean",
    .default = "other"
  )) %>%
  mutate(age_group = case_when(
    adult_age == 3 ~ 3,
    adult_age == 4 ~ 4,
    between(adult_age, 5, 6) ~ 5
  )) %>%
  # sort age
  arrange(age_group) %>%
  # set latent variable as ordered factor
  mutate(age_group = factor(age_group, ordered = TRUE))


# import model
m <- readRDS(here("results/models.local", "fit_DOY_npgo3.rds"))
```

```{r themeset, eval=FALSE, include=FALSE}
# https://bookdown.org/rdpeng/RProgDA/building-a-new-theme.html
# new_theme <- theme_minimal + theme(
#   axis.line.x = element_line(linewidth = 0.2),
#   axis.ticks = element_line(linewidth = 0.2),
#   axis.line.y = element_line(linewidth = 0.2),
#   text = element_text(
#     size = 12,
#     family = "Calibri"
#     ))

#theme_set(new_theme)
# ?additions to add from JG code:
# theme_minimal()+
# theme(strip.text = element_blank(),
#      axis.line.x = element_line(linewidth = 0.2),
#     axis.ticks = element_line(linewidth = 0.2),
#    axis.line.y = element_line(linewidth = 0.2),
#   legend.position = "top",
#  legend.key.size = unit(0.2, 'cm'),
# legend.text = element_text(size=7),
# axis.text.x = element_text(angle=90, vjust=.5, hjust = 1),
# axis.title.x = element_text(margin = margin(t=8, r=0, b=0, l=0)),
# axis.title.y = element_text(margin = margin(t=0, r=8, b=0, l=0)),
# plot.title = element_text(size = 10, vjust = -7))

# theme_set(theme_grey())#option to reset
# theme_get()
```

```{r theme_flextable, include=FALSE}
set_flextable_defaults(
  font.size = 12,
  font.family = "Cambria"
)
# font.color = "black",
# table.layout = "auto",
# padding.top = 3, padding.bottom = 3,
# padding.left = 4, padding.right = 4)
# theme_fun = "theme_box")

# init_flextable_defaults()
```


# Main Figures and Tables {-}

For pdf and png individual files see [github](https://github.com/ebuhle/chinook-adult-age/tree/master/results/figures/manuscript)


# Table 1. Covariates, sources, hypothesized mechanisms.

see manuscript

# Figure 2. Number of adults by ocean age with proportion inset
 
```{r sumreturns, include=FALSE}
# summarize data by return age
d.return <- data.scaled %>%
  mutate_if(sapply(data.scaled, is.character), as.factor) %>%
  group_by(pass_type_T_R, SYage1, OceanAge, OA.true) %>%
  summarise(n = n())
```



```{r p_proportions_inset, include=FALSE}
# Plot proportion inset figure

## plot
p.inset <- ggplot(d.return, aes(x = as.factor(SYage1), y = n, fill = forcats::fct_rev(OceanAge))) +
  geom_bar(position = position_fill(), stat = "identity", width = .8) +
  labs(
    x = "Smolt migration year", y = "Proportion",
    color = "",
    fill = "",
    title = NULL,
    subtitle = NULL,
    caption = ""
  ) +
  scale_x_discrete(breaks = every_nth(n = 2)) +
  ggtitle("(b) Proportion by age") +
  scale_fill_manual(
    values = pal_age,
    breaks = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    text = element_text(size = 8),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    plot.title = element_text(size = 8),
    panel.background = element_blank()
    )
```


```{r p_numberofreturns, fig.height=4, fig.width=7, dev= c('png', 'pdf')}
# facet
p.facet <- ggplot(d.return, aes(x = as.factor(SYage1), y = n)) +
  geom_bar(aes(fill = forcats::fct_rev(OceanAge)), position = position_stack(), stat = "identity", width = .8) +
  labs(
    x = "Smolt migration year", y = "Count of adult returns",
    color = NULL,
    fill = NULL,
    title = NULL,
    caption = ""
  ) +
  # geom_hline(aes(yintercept =mean(n)))+
  scale_fill_manual(
    values = pal_age,
    breaks = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  scale_x_discrete(breaks = every_nth(n = 2)) +
  ggtitle("(a) Count by age") +
  theme_minimal() +
  theme(
    strip.text = element_blank(),
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2),
    legend.position = "top",
    legend.key.size = unit(0.2, "cm"),
    legend.text = element_text(size = 7),
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1),
    axis.title.x = element_text(margin = margin(t = 8, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 8, b = 0, l = 0)),
    plot.title = element_text(size = 10, vjust = -7)
  ) +
  facet_grid(~OceanAge)

# place inset into facet figure
# pdf(here("results", "figures", "manuscript", "Fig1 Count by Age.pdf"), width=7, height=5, onefile=TRUE)
p.facet +
  patchwork::inset_element(p.inset, align_to = "plot", left = 0.76, bottom = 0.64, right = .99, top = 0.99)
```

# Figure 3. Beta parameters

## Figure 3 Beta parameters, reference level: MPG 
* see pdf version adjusted via ppt for a potential clustering option. Below is code for base figure.
```{r p_betaparameters, fig.width=8, fig.height=10, dev= c('png', 'pdf'),  cache=TRUE}
# could look into augmenting facet_grid/wrap as well to add color banners (issue is dropping unused variables), could make 3 plots and patchworks instead
# pdf(here("results", "figures", "manuscript", "Fig2 Beta params.pdf"), width=7, height=9, onefile=TRUE)
m %>%
  gather_draws(
    b_MPGLowerSnakeRiver, b_MPGClearwaterRiver,
    b_MPGSouthForkSalmonRiver, b_MPGLowerSalmonRiver,
    b_MPGMiddleForkSalmonRiver, b_MPGUpperSalmonRiver,
    b_length, b_rel_ageSmoltAboveLGR, b_rel_ageSmoltAtLGR,
    `b_length:rel_ageSmoltAboveLGR`, `b_length:rel_ageSmoltAtLGR`,
    b_SYage1_DOY, b_pass_type_T_RT, b_LGR.flow.7d, b_NPGO.ONDJFM.T.AVG3
  ) %>%
  mutate(subgroup = if_else(.variable %in% c("b_NPGO.ONDJFM.T.AVG3"), "Marine",
    if_else(.variable %in% c("b_pass_type_T_RT", "b_LGR.flow.7d"), "Freshwater", "Fish")
  )) %>%
  mutate(.variable = factor(.variable,
    levels = c(
      "b_NPGO.ONDJFM.T.AVG3",
      "b_LGR.flow.7d",
      "b_pass_type_T_RT",
      "b_SYage1_DOY",
      "b_MPGUpperSalmonRiver", "b_MPGMiddleForkSalmonRiver", "b_MPGLowerSalmonRiver",
      "b_MPGSouthForkSalmonRiver", "b_MPGClearwaterRiver", "b_MPGLowerSnakeRiver",
      "b_length:rel_ageSmoltAtLGR", "b_length:rel_ageSmoltAboveLGR",
      "b_rel_ageSmoltAtLGR", "b_rel_ageSmoltAboveLGR",
      "b_length"
    ),
    labels = c(
      "NPGO",
      "Flow",
      "Transported",
      "Day of Year",
      "Upper Salmon",
      "MF Salmon",
      "Lower Salmon",
      "SF Salmon",
      "Clearwater",
      "Lower Snake",
      "Length x \nSmolt&AtLGR",
      "Length x \nSmolt&AboveLGR",
      "Smolt&AtLGR",
      "Smolt&AboveLGR",
      "Length"
    )
  )) %>%
  ggplot(aes(y = .variable, x = .value)) + # reorder(variable, desc(.value))
  geom_vline(
    xintercept = 0, linetype = "dashed",
    color = "darkgrey"
  ) +
  stat_halfeye(
    aes(fill = subgroup, fill_ramp = after_stat(level)),
    .width = c(.50, .80, .90), scale = .7, # adjust based on CI wanted--poster includes .85 &.95 only
    # NOTE: we use position = "dodgejust" (a dodge that respects the
    # justification of intervals relative to slabs) instead of
    # position = "dodge" here because it ensures the topmost slab does
    # not extend beyond the plot limits
    position = "dodgejust"
  ) +
  # a range from 1 down to 0.2 ensures the fill goes dark to light inside-out
  # and doesn't get all the way down to white (0) on the lightest color
  scale_fill_ramp_discrete(na.translate = FALSE) +
  scale_x_continuous(breaks = seq(-1, 0.5, by = .5), limits = c(-1, 0.5)) +
  scale_y_discrete(expand = c(-.05, .9)) +
  scale_fill_manual(
    breaks = c("Fish", "Freshwater", "Marine"),
    values = pal_cov, guide = NULL
  ) +
  geom_hline(yintercept = 15.7, color = pal_cov[1], alpha = 0.3, linewidth = 3.5) +
  annotate("segment", x = -.98, y = 15.6, xend = -.98, yend = 3.83, color = pal_cov[1], alpha = .4, linewidth = 2.3) +
  annotate("text", y = 15.7, x = -0.98, label = "  Biological / behavioural covariates", color = "black", hjust = 0, size = 3.1) +
  
  geom_hline(yintercept = 3.7, color = pal_cov[2], alpha = 0.2, linewidth = 3.5) +
  annotate("segment", x = -.98, y = 3.6, xend = -.98, yend = 1.8, color = pal_cov[2], alpha = .3, linewidth = 2.3) +
  annotate("text", y = 3.7, x = -0.98, label = "  Freshwater covariates", color = "black", hjust = 0, size = 3.1) +
  
  geom_hline(yintercept = 1.7, color = pal_cov[3], alpha = 0.2, linewidth = 3.5) +
  annotate("segment", x = -.98, y = 1.6, xend = -.98, yend = 0.8, color = pal_cov[3], alpha = .3, linewidth = 2.3) +
  annotate("text", y = 1.7, x = -0.98, label = "  Marine covariates", color = "black", hjust = 0, size = 3.1) +
  
  annotate("segment", x = -.98, y = 15.1, xend = -.98, yend = 14.9, color = "#a37a29", alpha = .4, linewidth = 2.3) +
  
  annotate("text", x = -.775, y = 13.65, label = "Tag Age & Location", size = 3, color = "#a37a29") +
  annotate("text", x = -.73, y = 13.35, label = "Reference: Parr&AboveLGR", size = 2.3, color = "#a37a29") +
  annotate("segment", x = -.98, y = 14.3, xend = -.98, yend = 12.7, color = "#a37a29", alpha = .4, linewidth = 2.3) +
  
  annotate("text", x = -.69, y = 11.65, label = "Length x Tag Age & Location", size = 3, color = "#a37a29") +
  annotate("text", x = -.665, y = 11.35, label = "Reference: Length x Parr&AboveLGR", size = 2.3, color = "#a37a29") +
  annotate("segment", x = -.98, y = 12.3, xend = -.98, yend = 10.7, color = "#a37a29", alpha = .4, linewidth = 2.3) +
  
  annotate("text", x = -.9, y = 7.65, label = "Origin", size = 3, color = "#a37a29") +
  annotate("text", x = -.69, y = 7.35, label = "Reference: Grande Ronde/Imnaha", size = 2.3, color = "#a37a29") +
  annotate("segment", x = -.98, y = 10.3, xend = -.98, yend = 4.7, color = "#a37a29", alpha = .4, linewidth = 2.3) +
  annotate("segment", x = -.98, y = 4.1, xend = -.98, yend = 3.9, color = "#a37a29", alpha = .4, linewidth = 2.3) +
  
  annotate("text", x = -.7, y = 3.15, label = "Hydrosystem Passage Type", size = 3, color = pal_cov[2]) +
  annotate("text", x = -.8, y = 2.85, label = "Reference: In-river", size = 2.3, color = pal_cov[2]) +
  annotate("segment", x = -.98, y = 3.3, xend = -.98, yend = 2.7, color = pal_cov[2], alpha = .4, linewidth = 2.3) +
  annotate("segment", x = -.98, y = 2.1, xend = -.98, yend = 1.9, color = pal_cov[2], alpha = .4, linewidth = 2.3) +
  
  annotate("segment", x = -.98, y = 1.1, xend = -.98, yend = 0.9, color = pal_cov[3], alpha = .4, linewidth = 2.3) +
  labs(
    fill_ramp = "CI", # is this confidence or credible interval? double check
    fill = "Covariates",
    y = expression(bold("Covariate")),
    x = expression(bold(paste("Covariate effect (", beta, ")")))
  ) +
  theme_minimal() +
  theme(
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2, color = "gray"),
    axis.title.x = element_text(margin = margin(t = 8, r = 0, b = 0, l = 0)),
    legend.position = "right",
    legend.key.size = unit(0.3, "cm"),
    legend.text = element_text(size = 9),
    # axis.text.y = element_text(hjust = .01),
    plot.title.position = "plot"
  )
```

# Figure 4. Conditional effects


```{r condeffects, include = FALSE, echo=FALSE}
# categorical set
c_eff <- conditional_effects(m, prob = .9, categorical = TRUE, method = "posterior_epred", re_formula = NA) 

```


```{r p_individualcondeffects, cache=TRUE}

# length
FLz_xaxis <- (seq(40, 160, 10) - mean(DATA$length)) / sd(DATA$length)
FL_lab <- as.character(seq(40, 160, 10))

p_length <- plot(c_eff, plot = FALSE)[[1]] +
  scale_color_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  scale_shape_manual(guide = "none") +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  labs(x = "Length (mm)", y = "Probability", fill = NULL, color = NULL, shape = NULL, y = NULL) +
  ggtitle("(a) Length") +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.line.y = element_line(linewidth = 0.2),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    legend.key.size = unit(0.2, "cm"),
    legend.text = element_text(size = 9),
    plot.margin = unit(c(0.3, 0.3, 0.6, 0.3), "cm")
  ) +
  scale_x_continuous(breaks = FLz_xaxis, labels = FL_lab)


# Tag_Age&Location
p_rel_age <- plot(c_eff, plot = FALSE)[[2]] +
  scale_color_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater"),
    guide = "none"
  ) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater"),
    guide = "none"
  ) +
  scale_shape_manual(
    values = 1,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater"),
    guide = "none"
  ) +
  scale_x_discrete(
    breaks = c("ParrAboveLGR", "SmoltAboveLGR", "SmoltAtLGR"),
    labels = c("Parr&\nAboveLGR", "Smolt&\nAboveLGR", "Smolt&\nAtLGR")
  ) +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  labs(x = "Tag age & location", fill = NULL, color = NULL, shape = NULL, y = NULL) + # coord_capped_cart(bottom='both', left = "both", xlim= c(-3,3)) +
  ggtitle("(b) Tag age & location") +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2),
    # axis.text.x = element_text(angle=10),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    legend.key.size = unit(0.2, "cm"),
    legend.text = element_text(size = 9),
    plot.margin = unit(c(0.3, 0.3, 0.6, 0.3), "cm")
  )


# length x tag_age&location

c_eff_intrxn_1 <- conditional_effects(
  m,
  effects = "length",
  conditions = data.frame(rel_age = "ParrAboveLGR"),
  categorical = TRUE
)

p_intrxn_1 <- plot(c_eff_intrxn_1, plot = FALSE)[[1]] +
  scale_color_manual(values = c(
    "3" = pal_age[1],
    "4" = pal_age[2],
    "5" = pal_age[3]
  )) +
  scale_fill_manual(values = c(
    "3" = pal_age[1],
    "4" = pal_age[2],
    "5" = pal_age[3]
  )) +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  labs(x = "Length (mm)", y = "Probability", fill = NULL, color = NULL, shape = NULL) +
  ylab("Probability") +
  ggtitle("(c) incl. Length x ParrAboveLGR") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.line.y = element_line(linewidth = 0.2),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    plot.margin = unit(c(0.3, 0.3, 0.3, 0.3), "cm")
  ) +
  scale_x_continuous(breaks = FLz_xaxis, labels = FL_lab)


c_eff_intrxn_2 <- conditional_effects(
  m,
  effects = "length",
  conditions = data.frame(rel_age = "SmoltAboveLGR"),
  categorical = TRUE
)

p_intrxn_2 <- plot(c_eff_intrxn_2, plot = FALSE)[[1]] +
  scale_color_manual(values = c(
    "3" = pal_age[1],
    "4" = pal_age[2],
    "5" = pal_age[3]
  )) +
  scale_fill_manual(values = c(
    "3" = pal_age[1],
    "4" = pal_age[2],
    "5" = pal_age[3]
  )) +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  labs(x = "Length (mm)", y=NULL, fill = NULL, color = NULL, shape = NULL) +
  ggtitle("(d) incl. Length x SmoltAboveLGR") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.line.y = element_line(linewidth = 0.2),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    plot.margin = unit(c(0.3, 0.3, 0.3, 0.3), "cm")
  ) +
  scale_x_continuous(breaks = FLz_xaxis, labels = FL_lab)



c_eff_intrxn_3 <- conditional_effects(
  m,
  effects = "length",
  conditions = data.frame(rel_age = "SmoltAtLGR"),
  categorical = TRUE
)

p_intrxn_3 <- plot(c_eff_intrxn_3, plot = FALSE)[[1]] +
  scale_color_manual(values = c(
    "3" = pal_age[1],
    "4" = pal_age[2],
    "5" = pal_age[3]
  )) +
  scale_fill_manual(values = c(
    "3" = pal_age[1],
    "4" = pal_age[2],
    "5" = pal_age[3]
  )) +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  labs(x = "Length (mm)", y=NULL, fill = NULL, color = NULL, shape = NULL) +
  ggtitle("(e) incl. Length x SmoltAtLGR") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.line.y = element_line(linewidth = 0.2),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    plot.margin = unit(c(0.3, 0.3, 0.3, 0.3), "cm")
  ) +
  scale_x_continuous(breaks = FLz_xaxis, labels = FL_lab)



# Origin
p_origin <- plot(c_eff, plot = FALSE)[[3]] +
  scale_color_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater"),
    guide = "none"
  ) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater"),
    guide = "none"
  ) +
  scale_shape_manual(
    values = 1,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater"),
    guide = "none"
  ) +
  labs(x = "Origin", y = "Probability", fill = NULL, color = NULL, shape = NULL, y = NULL) +
  scale_x_discrete(
    breaks = c(
      "Grande Ronde/Imnaha", "Lower Snake River",
      "Clearwater River", "South Fork Salmon River",
      "Lower Salmon River", "Middle Fork Salmon River",
      "Upper Salmon River"
    ),
    labels = c(
      "Grande Ronde/\nImnaha", "L Snake",
      "Clearwater", "SF Salmon",
      "L Salmon", "MF Salmon",
      "U Salmon"
    )
  ) +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  ggtitle("(f) Fish tag & release origin") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    plot.margin = unit(c(0.3, 0.3, 0.6, 0.3), "cm")
  )
# axis.text.x = element_text(angle = 270,vjust=.5, hjust = 1)


# DOY
DOYz_xaxis <- (seq(80, 160, 10) - mean(DATA$SYage1_DOY)) / sd(DATA$SYage1_DOY)
DOY_lab <- as.character(seq(80, 160, 10))

p_DOY <- plot(c_eff, plot = FALSE)[[4]] +
  scale_color_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  scale_shape_manual(guide = "none") +
  labs(x = "Day of year", y = "Probability", fill = NULL, color = NULL, shape = NULL, y = NULL) +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  # scale_x_continuous(breaks = seq(-3,3,1))+
  ggtitle("(g) Day of year at LGR") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.line.y = element_line(linewidth = 0.2),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    plot.margin = unit(c(0.3, 0.3, 0.6, 0.3), "cm")
  ) +
  scale_x_continuous(breaks = DOYz_xaxis, labels = DOY_lab)


# pass_type
p_pass_type <- plot(c_eff, plot = FALSE)[[6]] +
  scale_color_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater"),
    guide = "none"
  ) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater"),
    guide = "none"
  ) +
  scale_shape_manual(
    values = 1,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater"),
    guide = "none"
  ) +
  scale_x_discrete(labels = c("In-river", "Transported")) +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  labs(x = "Passage type", fill = NULL, color = NULL, shape = NULL, y = NULL) +
  ggtitle("(h) Hydrosystem passage type") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    plot.margin = unit(c(0.3, 0.3, 0.6, 0.3), "cm")
  )



# flow
flowz_xaxis <- (seq(0, 300, 20) - mean(DATA$LGR.flow.7d)) / sd(DATA$LGR.flow.7d)
flow_lab <- as.character(seq(0, 300, 20))

p_flow <- plot(c_eff, plot = FALSE)[[5]] +
  scale_color_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  scale_shape_manual(guide = "none") +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  labs(x = "Flow (kcfs)", y = "Probability", fill = NULL, color = NULL, shape = NULL, y = NULL) +
  ggtitle("(i) Flow at LGR") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.line.y = element_line(linewidth = 0.2),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    plot.margin = unit(c(0.3, 0.3, 0.3, 0.3), "cm")
  ) +
  scale_x_continuous(breaks = flowz_xaxis, labels = flow_lab)


# NPGO
npgoz_xaxis <- (seq(-2.5, 2.5, 0.5) - mean(DATA$NPGO.ONDJFM.T)) / sd(DATA$NPGO.ONDJFM.T)
npgo_lab <- as.character(seq(-2.5, 2.5, 0.5))

p_NPGO <- plot(c_eff, plot = FALSE)[[7]] +
  scale_color_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("3", "4", "5"),
    labels = c("1-Ocean", "2-Ocean", "3-Ocean or greater")
  ) +
  scale_shape_manual(guide = "none") +
  scale_y_continuous(breaks = seq(0, 0.8, by = .2), limits = c(0, 0.8)) +
  labs(x = "NPGO index", fill = NULL, color = NULL, shape = NULL, y = NULL) +
  ggtitle("(j) NPGO index, 3-year ONDJFM mean") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 9.5),
    plot.margin = unit(c(0.3, 0.3, 0.3, 0.3), "cm")
  ) +
  scale_x_continuous(breaks = npgoz_xaxis, labels = npgo_lab)

```

## Figure 4 Conditional effects, (a)-(e)

```{r p_condeffects_a-f, echo=FALSE, fig.height=8, fig.width=7, dev= c('png', 'pdf'),  cache=TRUE}
# patch1<-p8 / (p2 + p7) / (p9+plot_layout(guides = "keep")) / (p1 + p3) + plot_layout(guides = "collect") & ylab(NULL) & theme_classic() & theme(legend.position = "bottom")

# wrap_elements(patch1) +
#   labs(tag = "Probability") +
#  theme(
#    plot.tag = element_text(size = rel(1), angle = 90),
#    plot.tag.position = "left"
#  )


# pdf(here("results", "figures", "manuscript", "Fig4a_e length effects.pdf"), width=8, height=7, onefile=TRUE)
patch1 <- (p_length + p_rel_age) / (p_intrxn_1 + p_intrxn_2 + p_intrxn_3)
patch1
```

## Figure 4 Conditional effects, (f)-(i)

```{r p_condeffects_g-i, echo=FALSE, fig.height=8, fig.width=7, dev= c('png', 'pdf'),  cache=TRUE}
# patch2<-(p4 + p5 + plot_layout(ncol=2))+(p6 + plot_layout(ncol=2))+ plot_layout(guides = "collect") & ylab(NULL) & theme_classic() + theme( legend.position = "bottom")


# wrap_elements(patch2) +
#   labs(tag = "Probability") +
#   theme(
#     plot.tag = element_text(size = rel(1), angle = 90),
#     plot.tag.position = "left"
#   )

# pdf(here("results", "figures", "manuscript", "Fig4f_j cond effects.pdf"), width=7, height=9, onefile=TRUE)
patch2 <- p_origin / (p_DOY + p_pass_type ) / (p_flow + p_NPGO)
patch2
```

# Figure 5. Model fit


## Figure 5a. Model fit: proportion

```{r  p_modelfit_proportion, echo=FALSE, fig.height=9, fig.width=8,dev= c('png', 'pdf'),  cache=TRUE}
# using ppc_check (preference)
# set Y
# pdf(here("results", "figures", "manuscript", "Fig5a Obs vs Pred proportions no RE.pdf"), width=7, height=9, onefile=TRUE)
y_int <- as.integer(data.scaled$age_group) 

# set yrep
yrep_int <- posterior_predict(m, newdata = NULL, re_formula = NA)

color_scheme_set(scheme = "mix-darkgray-gray")


ppc_bars_grouped(
  y = y_int,
  yrep = yrep_int,
  group = data.scaled$SYage1,
  freq = FALSE, # adjusts from TRUE = count and FALSE = proportion
  prob = 0.9,
  fatten = 1,
  size = 1.5,
  facet_args = list(
    ncol = 3,
    scales = "fixed"
  ) # "fixed" or "free_y"
) + scale_x_continuous(
  breaks = c(1, 2, 3), # check back on this
  labels = c("1-Ocean", "2-Ocean", "3-Ocean")
) +
  labs(x = "Ocean Age") + theme(text = element_text(size = 14)) +
  theme_tidybayes() +
  theme(
    legend.position = c(.85, .05),
    legend.key.size = unit(.5, "cm"),
    legend.spacing.y = unit(.02, "cm"),
    legend.background = element_blank()
  )
```

## Figure 5b. Model fit: count
```{r  p_modelfit_count, echo=FALSE, fig.height=9, fig.width=8,dev= c('png', 'pdf'),  cache=TRUE}
# pdf(here("results", "figures", "manuscript", "Fig5b Obs vs Pred counts no RE.pdf"), width=7, height=9, onefile=TRUE)

ppc_bars_grouped(
  y = y_int,
  yrep = yrep_int,
  group = data.scaled$SYage1,
  freq = TRUE, # adjusts from TRUE = count and FALSE = proportion
  prob = 0.9,
  fatten = 1,
  size = 1.5,
  facet_args = list(
    ncol = 3,
    scales = "free_y"
  ) # "fixed" or "free_y"
) + scale_x_continuous(
  breaks = c(1, 2, 3),
  labels = c("1-Ocean", "2-Ocean", "3-Ocean")
) +
  labs(x = "Ocean Age", y = "Count of adult returns") + theme(text = element_text(size = 14)) +
  theme_tidybayes() +
  theme(
    legend.position = "none",
    strip.text.x = element_text(margin = margin(0.05, 0, 0.05, 0, "cm")),
    strip.text = element_text(size = 8),
    panel.spacing = unit(1, "lines")
  )
# theme(legend.position = c(.85,.05),
#                                  legend.key.size = unit(.5, "cm"),
#                                  legend.spacing.y = unit(.02, 'cm'),
#                                  legend.background = element_blank())
```


## Figure 5c. Model fit: proportion

```{r  p_modelfit_proportion, echo=FALSE, fig.height=9, fig.width=8,dev= c('png', 'pdf'),  cache=TRUE}
# using ppc_check (preference)
# set Y
# pdf(here("results", "figures", "manuscript", "Fig5c Obs vs Pred proportions no RE.pdf"), width=7, height=9, onefile=TRUE)
y_int <- as.integer(data.scaled$age_group) 

# set yrep
yrep_int <- posterior_predict(m, newdata = NULL, re_formula = NA)

color_scheme_set(scheme = "mix-darkgray-gray")


ppc_bars_grouped(
  y = y_int,
  yrep = yrep_int,
  group = data.scaled$SYage1,
  freq = FALSE, # adjusts from TRUE = count and FALSE = proportion
  prob = 0.95,
  fatten = 1,
  size = 1.5,
  facet_args = list(
    ncol = 3,
    scales = "fixed"
  ) # "fixed" or "free_y"
) + scale_x_continuous(
  breaks = c(1, 2, 3), # check back on this
  labels = c("1-Ocean", "2-Ocean", "3-Ocean")
) +
  labs(x = "Ocean Age") + theme(text = element_text(size = 14)) +
  theme_tidybayes() +
  theme(
    legend.position = c(.85, .05),
    legend.key.size = unit(.5, "cm"),
    legend.spacing.y = unit(.02, "cm"),
    legend.background = element_blank()
  )
```

## Figure 5d. Model fit: count, no random effects
```{r  p_modelfit_count, echo=FALSE, fig.height=9, fig.width=8,dev= c('png', 'pdf'),  cache=TRUE}
# pdf(here("results", "figures", "manuscript", "Fig5d Obs vs Pred counts no RE.pdf"), width=7, height=9, onefile=TRUE)

ppc_bars_grouped(
  y = y_int,
  yrep = yrep_int,
  group = data.scaled$SYage1,
  freq = TRUE, # adjusts from TRUE = count and FALSE = proportion
  prob = 0.95,
  fatten = 1,
  size = 1.5,
  facet_args = list(
    ncol = 3,
    scales = "fixed"
  ) # "fixed" or "free_y"
) + scale_x_continuous(
  breaks = c(1, 2, 3),
  labels = c("1-Ocean", "2-Ocean", "3-Ocean")
) +
  labs(x = "Ocean Age", y = "Count of adult returns") + theme(text = element_text(size = 14)) +
  theme_tidybayes() +
  theme(
    legend.position = "none",
    strip.text.x = element_text(margin = margin(0.05, 0, 0.05, 0, "cm")),
    strip.text = element_text(size = 8),
    panel.spacing = unit(1, "lines")
  )
# theme(legend.position = c(.85,.05),
#                                  legend.key.size = unit(.5, "cm"),
#                                  legend.spacing.y = unit(.02, 'cm'),
#                                  legend.background = element_blank())
```

# Supplementary Figures {-}


# Table S1. Sample size
## Table S1a. Number of adult returns by passage type and ocean age per smolt migration year
```{r t_samplesize_bypassagetype, echo=FALSE}
tb1 <- DATA %>%
  group_by(SYage1, pass_type_T_R, adult_age) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = c(pass_type_T_R, adult_age), values_from = count) %>%
  select(SYage1, "ROR_3", "ROR_4", "ROR_5", "ROR_6", "T_3", "T_4", "T_5", "T_6") %>%
  replace(is.na(.), 0)


tb2 <- DATA %>%
  group_by(SYage1, pass_type_T_R) %>%
  summarise(sum = sum(n())) %>%
  pivot_wider(names_from = pass_type_T_R, values_from = sum)

tj <- full_join(tb1, tb2, by = "SYage1") %>%
  select(SYage1, "ROR_3", "ROR_4", "ROR_5", "ROR_6", "ROR", "T_3", "T_4", "T_5", "T_6", "T") %>%
  rename(
    "Juvenile\nmigration\nyear" = SYage1,
    "ROR_n" = ROR,
    "T_n" = T
  )


f.tj <- flextable(tj) %>%
  span_header() %>%
  labelizor(
    part = "header",
    labels = c(
      "ROR" = "In-River",
      "T" = "Transported"
    )
  ) %>%
  align(align = "left", part = "all") %>%
  font(fontname = "Calibri") %>%
  colformat_num(
    j = "Juvenile\nmigration\nyear",
    big.mark = ""
  ) %>%
  bold(j = c(6, 11), bold = TRUE, part = "all") %>%
  autofit()


f.tj

# print(f.tj, preview = "docx")
```


## Table S1b. Number of adult returns by origin and ocean age per smolt migration year
```{r t_samplesize_byMPG, echo=FALSE}
# need to add zero lines for sites with no data
tb1 <- DATA %>%
  group_by(MPG, adult_age) %>%
  summarise(count = n()) %>%
  pivot_wider(
    names_from = adult_age, values_from = count,
    values_fill = 0, names_sort = TRUE,
    names_expand = TRUE
  )

tb2 <- DATA %>%
  group_by(MPG) %>%
  summarise(n = sum(n()))

tj <- full_join(tb1, tb2, by = c("MPG")) %>%
  mutate(MPG = factor(MPG,
    levels = c(
      "Grande Ronde/Imnaha", "Lower Snake River", "Clearwater River",
      "Lower Salmon River", "South Fork Salmon River", "Middle Fork Salmon River",
      "Upper Salmon River"
    ),
    labels = c(
      "Grande Ronde Imnaha River", "Lower Snake River", "Clearwater River",
      "Lower Salmon River", "South Fork Salmon River", "Middle Fork Salmon River",
      "Upper Salmon River"
    )
  ))

f.tj <- flextable(tj)


f.adj <- f.tj %>%
  labelizor(
    part = "header",
    labels = c("MPG" = "Origin")
  ) %>%
  add_header_row(
    values = c("", "Ocean age", ""),
    colwidths = c(1, 4, 1), top = TRUE
  ) %>%
  bold(j = c(6), bold = TRUE, part = "all") %>%
  align(align = "left", part = "all") %>%
  font(fontname = "Calibri") %>%
  hline(part = "header") %>%
  autofit()

f.final <- set_table_properties(
  f.adj,
  layout = "autofit",
  width = 0,
  align = "center",
  word_title = NULL,
  word_description = NULL
)
f.final

# print(f.final, preview = "docx")

# will need to either figure out how to manipulate header (ROR/T) using separate_header_label ( was able to do this to row 2 header, just not row 1 headers) or go longer route which just adds generic header.
```


## Table S1c. Number of adult returns by tag_age&location and ocean age per smolt migration year
```{r t_samplesize_byrelage, echo=FALSE}
tb1 <- DATA %>%
  group_by(SYage1, rel_age, adult_age) %>%
  summarise(count = n()) %>%
  pivot_wider(
    names_from = c(rel_age, adult_age), values_from = count,
    values_fill = 0, names_sort = TRUE,
    names_expand = TRUE
  )

tb2 <- DATA %>%
  group_by(SYage1, rel_age) %>%
  summarise(sum = sum(n())) %>%
  pivot_wider(
    names_from = rel_age, values_from = sum,
    values_fill = 0, names_glue = "{rel_age}_n",
    names_sort = TRUE, names_expand = TRUE
  )

tj <- full_join(tb1, tb2, by = "SYage1") %>%
  select(SYage1, contains("parr"), contains("SmoltAbove"), contains("SmoltAt"))



f.tj <- flextable(tj) %>%
  span_header()


f.adj <- labelizor(
  x = f.tj,
  part = "header",
  labels = c(
    "SYage1" = "Juvenile\nmigration\nyear",
    "ParrAboveLGR" = "Parr&AboveLGR",
    "SmoltAboveLGR" = "Smolt&AboveLGR",
    "SmoltAtLGR" = "Smolt&AtLGR"
  )
)

f.adj <- colformat_num(f.adj,
  j = 1,
  big.mark = ""
) %>%
  bold(j = c(6, 11, 16), bold = TRUE, part = "all") %>%
  align(align = "left", part = "all") %>%
  font(fontname = "Calibri") %>%
  hline(part = "header") %>%
  autofit()


f.final <- set_table_properties(
  f.adj,
  layout = "autofit",
  width = 0,
  align = "center",
  word_title = NULL,
  word_description = NULL
)
f.final

# print(f.adj, preview = "docx")
```





## Figure S1a. Total number of adult returns by ocean age
```{r p_totalnumberofreturns, fig.height=4, fig.width=6, dev= c('png', 'pdf')}
# facet
ggplot(d.return, aes(x = as.factor(SYage1), y = n)) +
  geom_bar(aes(fill = forcats::fct_rev(OA.true)), position = position_stack(), stat = "identity", width = .6) +
  labs(
    x = "Smolt migration year", y = "Number of adult returns",
    color = NULL,
    fill = NULL
  ) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("1-Ocean", "2-Ocean", "3-Ocean", "4-Ocean")
  ) +
  scale_x_discrete(breaks = every_nth(n = 1)) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  )
```


# Environmental covariates vs year {-}
# Juvenile covariates{-}


```{r, eval=FALSE, include=FALSE}
# plot to cut off at data
ggplot(DATA, aes(y = fct_rev(as.factor(SYage1)), x = length)) +
  geom_density_ridges(aes(height = ..density..), color = "white", fill = "darkgrey", scale = 1, stat = "density", trim = TRUE) + # scale = 1
  geom_point(aes(x = length), color = "white", fill = "darkgrey", shape = 21, size = .5, alpha = .75, position = position_nudge(x = -.09)) +
  geom_vline(xintercept = median(DATA$length), linetype = "solid", color = "darkgrey", size = .3) +
  geom_vline(xintercept = max(DATA$length), linetype = "dashed", color = "darkgrey", size = .3) +
  geom_vline(xintercept = min(DATA$length), linetype = "dashed", color = "darkgrey", size = .3) +
  labs(
    x = "Fork Length (mm)",
    y = "Smolt migration year",
    fill = NULL
  ) +
  # coord_cartesian(clip = "off") +
  theme(
    legend.position = "none",
    axis.line.x = element_line(color = "black")
  )

### or use stat_slab() options
# if we want to highlight 95%, but extends ridgeline
ggplot(DATA, aes(y = fct_rev(as.factor(SYage1)), x = LGR.flow.7d)) +
  stat_slab(p_limits = c(0.05, 0.95), trim = T) +
  scale_fill_manual(values = c("lightgrey", "darkgrey", "lightgrey")) +
  geom_point(aes(x = LGR.flow.7d), color = "white", fill = "darkgrey", shape = 21, size = .5, alpha = .75, position = position_nudge(x = -.09), stroke = .25) +
  geom_vline(xintercept = median(DATA$LGR.flow.7d), linetype = "solid", color = "black", size = .15) +
  geom_vline(xintercept = max(DATA$LGR.flow.7d), linetype = "dashed", color = "black", size = .1) +
  geom_vline(xintercept = min(DATA$LGR.flow.7d), linetype = "dashed", color = "black", size = .1) +
  labs(
    x = "LGR temperature (°C)",
    y = "Smolt migration year",
    fill = NULL
  ) +
  scale_x_continuous(breaks = seq(20, 200, 20), limits = c(20, 200)) +
  # coord_cartesian(clip = "off") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(color = "white", size = .1),
    axis.line.y = element_line(color = "white", size = .1),
    axis.ticks = element_blank(),
    panel.grid.major.y = element_line(color = "lightgrey", size = .1),
    panel.grid.major.x = element_line(color = "lightgrey", size = .1),
    text = element_text(family = "Arial")
  )

# option deleted could use for all continuous?
# if we want to highlight 95%, but extends ridgeline
ggplot(DATA, aes(y = fct_rev(as.factor(SYage1)), x = SYage1_DOY)) +
  stat_density_ridges(aes(fill = stat(quantile)),
    quantile_lines = F,
    calc_ecdf = TRUE,
    geom = "density_ridges_gradient",
    quantiles = c(0.05, 0.95),
    scale = .9,
    color = "white",
    lwd = .25,
    rel_min_height = .05
  ) +
  scale_fill_manual(values = c("lightgrey", "darkgrey", "lightgrey")) +
  geom_point(aes(x = SYage1_DOY), color = "white", fill = "darkgrey", shape = 21, size = .5, alpha = .75, position = position_nudge(x = -.09), stroke = .25) +
  geom_vline(xintercept = median(DATA$SYage1_DOY), linetype = "solid", color = "black", size = .15) +
  geom_vline(xintercept = max(DATA$SYage1_DOY), linetype = "dashed", color = "black", size = .1) +
  geom_vline(xintercept = min(DATA$SYage1_DOY), linetype = "dashed", color = "black", size = .1) +
  labs(
    x = "LGR passage (DOY)",
    y = "Smolt migration year",
    fill = NULL
  ) +
  scale_x_continuous(breaks = seq(80, 160, 20), limits = c(80, 160)) +
  # coord_cartesian(clip = "off") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(color = "white", size = .1),
    axis.line.y = element_line(color = "white", size = .1),
    axis.ticks = element_blank(),
    panel.grid.major.y = element_line(color = "lightgrey", size = .1),
    panel.grid.major.x = element_line(color = "lightgrey", size = .1),
    text = element_text(family = "Arial")
  )
```

## Figure S2a. passage type

```{r pct_transport,include=FALSE}
df.sum <- data.scaled %>%
  group_by(SYage1) %>%
  summarise(n = n())

df.t <- data.scaled %>%
  group_by(SYage1, pass_type_T_R) %>%
  filter(pass_type_T_R == "T") %>%
  summarise(nT = n())

df.pctT <- inner_join(df.t, df.sum, by = "SYage1") %>%
  mutate(pctT = scales::percent(nT / n)) %>%
  mutate(pctR = scales::percent(1-(nT/n)))
```

```{r p_categoricalcovariatesbypassagetype,fig.height= 4, fig.width=7,dev= c('png', 'pdf'),  cache=TRUE}
# passage type
ggplot(d.return, aes(x = as.factor(SYage1))) +
  geom_bar(aes(fill = pass_type_T_R, y = n), position = "stack", stat = "identity", width = .6 ) +
  geom_text(data = df.pctT, aes(label = pctT, y = n, group = SYage1), position = "identity", vjust = -.5, hjust = .3, size = 3) +
  scale_fill_grey(
    breaks = c("ROR", "T"),
    labels = c("In-river", "Transported"),
    start = 0.8,
    end = 0.2
  ) +
  labs(
    x = "Smolt migration year", y = "Number of adult returns",
    fill = NULL
  ) +
  theme_classic() +
  theme(
    legend.position = c(.9, .8),
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) 
```



## Figure S2b. Number of adult returns by passage type and ocean age

```{r p_numberofreturnsbypassagetype, fig.height=4, fig.width=7, dev= c('png', 'pdf'), }
# New facet label names for transport
p.labs <- c("In-river", "Transported")
names(p.labs) <- c("ROR", "T")

# facet by transport
ggplot(d.return, aes(x = as.factor(SYage1), y = n, fill = forcats::fct_rev(OA.true))) +
  geom_bar(position = "stack", stat = "identity", width = .8) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("1-Ocean", "2-Ocean", "3-Ocean", "4-Ocean")
  ) +
  labs(
    x = "Smolt migration year", y = "Number of adult returns",
    color = NULL,
    fill = NULL
  ) +
  scale_x_discrete(breaks = every_nth(n = 1)) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
  facet_wrap(~pass_type_T_R, labeller = labeller(pass_type_T_R = p.labs))
```


## Figure S3a. tag_age&location grouping
```{r p_categoricalcovariatesbyrelage,fig.height= 4, fig.width=7,dev= c('png', 'pdf'),  cache=TRUE}
# release age
DATA %>%
  group_by(SYage1, rel_age) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = as.factor(SYage1), y = n, fill = rel_age)) +
  geom_bar(position = "stack", stat = "identity", width = .6) +
  scale_fill_grey(
    breaks = c("ParrAboveLGR", "SmoltAboveLGR", "SmoltAtLGR"),
    labels = c("Parr&AboveLGR", "Smolt&AboveLGR", "Smolt&AtLGR")
  ) +
  labs(
    x = "Smolt migration year", y = "Number of adult returns",
    color = NULL,
    fill = NULL
  ) +
  theme_classic() +
  theme(
    legend.position = c(.85, .8),
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  )

```

## Figure S3b. tag_age&location grouping by passage type
```{r p_categoricalcovariates_releaseagelocation_bypassagetype, fig.height= 4, fig.width=7,dev= c('png', 'pdf'),  cache=TRUE}
# release age
DATA %>%
  group_by(SYage1, OceanAge, pass_type_T_R, rel_age) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = as.factor(SYage1), y = n, fill = rel_age)) +
  geom_bar(position = "stack", stat = "identity", width = .6) +
  scale_fill_grey(
    breaks = c("ParrAboveLGR", "SmoltAboveLGR", "SmoltAtLGR"),
    labels = c("Parr/aboveLGR", "Smolt/aboveLGR", "Smolt/atLGR")
  ) +
  labs(
    x = "Smolt migration year", y = "Number of adult returns",
    color = NULL,
    fill = NULL
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
  facet_wrap(~pass_type_T_R, labeller = labeller(pass_type_T_R = p.labs))
```


## Figure S4a. proportion of origin groupingg: steam plot
```{r  p_categoricalcovariates_releaseMPG_steam, fig.height= 4, fig.width=7,dev= c('png', 'pdf'),  cache=TRUE}
library(readxl)
library(ggstream)
library(showtext)
library(ggtext)


# Stacked area chart with smoothing via steam plot
DATA %>%
  group_by(SYage1, MPG) %>%
  summarise(n = n()) %>%
  mutate(MPG = factor(MPG, levels = c(
    "Grande Ronde/Imnaha", "Lower Snake River",
    "Clearwater River", "South Fork Salmon River",
    "Lower Salmon River", "Middle Fork Salmon River",
    "Upper Salmon River"
  ))) %>%
  ggplot(aes(x = SYage1, y = n, fill = MPG, label = MPG)) +
  geom_stream(type = "proportion") +
  labs(
    fill = "Origin",
    x = "Smolt migration year",
    y = "Proportion"
  ) +
  scale_x_continuous(breaks = seq(1998, 2020, by = 2)) +
  scale_fill_manual(
    breaks = c(
      "Grande Ronde/Imnaha", "Lower Snake River",
      "Clearwater River", "South Fork Salmon River",
      "Lower Salmon River", "Middle Fork Salmon River",
      "Upper Salmon River"
    ),
    labels = c(
      "Grande Ronde/Imnaha", "L Snake",
      "Clearwater", "SF Salmon",
      "L Salmon", "MF Salmon",
      "U Salmon"
    ),
    values = pal_all
  ) +
  theme_classic()
```

OR

## Figure S4a. proportion of origin grouping: stacked barplot
```{r  p_categoricalcovariates_releaseMPG_bar, fig.height= 4, fig.width=7,dev= c('png', 'pdf'),  cache=TRUE}
# bar
DATA %>%
  group_by(SYage1, MPG) %>%
  summarise(n = n()) %>%
  mutate(MPG = factor(MPG, levels = c(
    "Grande Ronde/Imnaha", "Lower Snake River", # look into ordering data with fill order
    "Clearwater River", "South Fork Salmon River",
    "Lower Salmon River", "Middle Fork Salmon River",
    "Upper Salmon River"
  ))) %>%
  ggplot(aes(x = SYage1, y = n, fill = MPG, label = MPG)) +
  geom_bar(position = position_fill(), stat = "identity", width = .8) +
  labs(
    fill = "Origin",
    x = "Smolt migration year",
    y = "Proportion"
  ) +
  scale_x_continuous(breaks = seq(1998, 2020, by = 1)) +
  scale_fill_manual(
    breaks = c(
      "Grande Ronde/Imnaha", "Lower Snake River", # look into ordering data with fill order
      "Clearwater River", "South Fork Salmon River",
      "Lower Salmon River", "Middle Fork Salmon River",
      "Upper Salmon River"
    ),
    labels = c(
      "Grande Ronde/Imnaha", "L Snake",
      "Clearwater", "SF Salmon",
      "L Salmon", "MF Salmon",
      "U Salmon"
    ),
    values = pal_all
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
```


## Figure S4b. Total number of origin grouping
```{r p_categoricalcovariatesbyreleaseMPG,fig.height= 4, fig.width=7,dev= c('png', 'pdf'),  cache=TRUE}
# release MPG index
DATA %>%
  group_by(SYage1, OceanAge, MPG) %>%
  summarise(n = n()) %>%
  mutate(MPG = factor(MPG, levels = c(
    "Grande Ronde/Imnaha", "Lower Snake River",
    "Clearwater River", "South Fork Salmon River",
    "Lower Salmon River", "Middle Fork Salmon River",
    "Upper Salmon River"
  ))) %>%
  ggplot(aes(x = as.factor(SYage1), y = n, fill = MPG)) +
  geom_bar(position = "stack", stat = "identity", width = .6) +
  scale_fill_manual(
    breaks = c(
      "Grande Ronde/Imnaha", "Lower Snake River",
      "Clearwater River", "South Fork Salmon River",
      "Lower Salmon River", "Middle Fork Salmon River",
      "Upper Salmon River"
    ),
    labels = c(
      "Grande Ronde/Imnaha", "L Snake",
      "Clearwater", "SF Salmon",
      "L Salmon", "MF Salmon",
      "U Salmon"
    ),
    values = pal_all
  ) +
  labs(
    x = "Smolt migration year", y = "Number of adult returns",
    color = NULL,
    fill = NULL
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 2)) #+ facet_wrap(~OceanAge)
```




## Figure S4c. Number of origin grouping by passage type

```{r p_categoricalcovariates_releaseMPG_bypassagetype, fig.height= 4, fig.width=7,dev= c('png', 'pdf'),  cache=TRUE}
# release MPG index
DATA %>%
  group_by(SYage1, OceanAge, pass_type_T_R, MPG) %>%
  summarise(n = n()) %>%
  mutate(MPG = factor(MPG, levels = c(
    "Grande Ronde/Imnaha", "Lower Snake River",
    "Clearwater River", "South Fork Salmon River",
    "Lower Salmon River", "Middle Fork Salmon River",
    "Upper Salmon River"
  ))) %>%
  ggplot(aes(x = as.factor(SYage1), y = n, fill = MPG)) +
  geom_bar(position = "stack", stat = "identity", width = .6) +
  scale_fill_manual(
    breaks = c(
      "Grande Ronde/Imnaha", "Lower Snake River",
      "Clearwater River", "South Fork Salmon River",
      "Lower Salmon River", "Middle Fork Salmon River",
      "Upper Salmon River"
    ),
    labels = c(
      "Grande Ronde/Imnaha", "L Snake",
      "Clearwater", "SF Salmon",
      "L Salmon", "MF Salmon",
      "U Salmon"
    ),
    values = pal_all
  ) +
  labs(
    x = "Smolt migration year", y = "Number of adult returns",
    color = NULL,
    fill = NULL
  ) +
  theme_classic() +
  # scale_x_discrete(breaks = every_nth(n = 2))+
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 2)) +
  facet_wrap(~pass_type_T_R, labeller = labeller(pass_type_T_R = p.labs))
```







## Figure S4d. Number of adult returns by origin and ocean age vy smolt migration year

```{r p_numberofreturnsbyorigin, fig.height=6, fig.width=6, dev= c('png', 'pdf')}
# summarize data by return age
data.scaled %>%
  mutate_if(sapply(data.scaled, is.character), as.factor) %>%
  group_by(SYage1, MPG, OceanAge, OA.true) %>%
  summarise(n = n()) %>%
  mutate(MPG = factor(MPG,
    levels = c(
      "Grande Ronde/Imnaha", "Lower Snake River",
      "Clearwater River", "South Fork Salmon River",
      "Lower Salmon River", "Middle Fork Salmon River",
      "Upper Salmon River"
    ),
    labels = c(
      "Grande Ronde/Imnaha", "L Snake",
      "Clearwater", "SF Salmon",
      "L Salmon", "MF Salmon",
      "U Salmon"
    )
  )) %>%
  # facet by transport
  ggplot(aes(x = as.factor(SYage1), y = n, fill = forcats::fct_rev(OA.true))) +
  geom_bar(position = "stack", stat = "identity", width = .8) +
  scale_fill_manual(
    values = pal_age,
    breaks = c("1-Ocean", "2-Ocean", "3-Ocean", "4-Ocean"),
    guide = guide_legend(nrow = 2)
  ) +
  labs(
    x = "Smolt migration year", y = "Number of adult returns",
    color = NULL,
    fill = "Ocean age"
  ) +
  scale_x_discrete(breaks = every_nth(n = 1)) +
  theme_classic() +
  facet_wrap(~MPG, ncol = 2) +
  theme(
    legend.position = c(.75, .05),
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  )
```





### Figure S5. Fork length

```{r p_continuouscovariates_forklength_a, error=FALSE, warning=FALSE, messages = FALSE, fig.height=6, fig.width=4}
# pdf(here("results", "figures", "manuscript", "FigS5 Fork length.pdf"), width=4, height=9, onefile=TRUE)

FLz_xaxis <- (seq(40, 160, 10) - mean(DATA$length)) / sd(DATA$length)
FL_lab <- as.character(seq(40, 160, 10))

ggplot(data.scaled, aes(x = length, y = as.character(SYage1), fill = factor(after_stat(quantile)))) +
  stat_density_ridges(
    rel_min_height = 0.03, scale = 0.9,
    quantile_lines = TRUE, quantiles = c(0.025, 0.5, 0.975),
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    jittered_points = TRUE,
    position = position_points_jitter(width = 0.05, height = 0),
    point_shape = "o", point_size = 0.7, point_alpha = 0.7, alpha = 0.5, size = .25
  ) +
  scale_fill_manual(
    name = "Probability", values = alpha(c("#333333", "#CCCCCC", "#CCCCCC", "#333333"), 0.8),
    labels = c("(0, 0.025]", "(0.025, 0.5]", "(0.5, 0.975]", "(0.975, 1]")
  ) +
  labs(x = "Fork Length (mm)", y = "Smolt Migration Year") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2)
  ) +
  scale_x_continuous(breaks = FLz_xaxis, labels = FL_lab)
```


```{r p_continuouscovariates_forklength,fig.height= 4, fig.width=6,dev= c('png', 'pdf'), fig.cap=" Dashed line represents median value: 109 mm",  cache=TRUE}
ggplot(DATA, aes(x = as.factor(SYage1), y = length), group = as.factor(SYage1)) +
  geom_point(shape = 21, fill = "transparent", color = "#ABABAB", stroke = .1, size = 1) +
  geom_boxplot(
    fill = "transparent", outlier.fill = "transparent",
    outlier.color = "#ABABAB", outlier.stroke = .1,
    outlier.size = 1, outlier.shape = 21
  ) +
  labs(
    y = "Length (mm)",
    x = "Smolt migration year",
    fill = NULL
  ) +
  geom_hline(yintercept = median(DATA$length), color = "black", linetype = "dashed") +
  scale_y_continuous(breaks = seq(40, 160, 20), limits = c(40, 160)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
```

## Figure S6a. DOY


```{r p_continuouscovariates_DOY_a,fig.height= 6,fig.width=4, dev= c('png', 'pdf'),fig.cap="Dashed line represents median value: 122 (DOY)",  cache=TRUE}
# pdf(here("results", "figures", "manuscript", "FigS6a DOY.pdf"), width=4, height=9, onefile=TRUE)

DOYz_xaxis <- (seq(80, 160, 10) - mean(DATA$SYage1_DOY)) / sd(DATA$SYage1_DOY)
DOY_lab <- as.character(seq(80, 160, 10))

ggplot(data.scaled, aes(x = SYage1_DOY, y = as.character(SYage1), fill = factor(after_stat(quantile)))) +
  stat_density_ridges(
    rel_min_height = 0.03, scale = 0.9,
    quantile_lines = TRUE, quantiles = c(0.025, 0.5, 0.975),
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    jittered_points = TRUE,
    position = position_points_jitter(width = 0.05, height = 0),
    point_shape = "o", point_size = 0.7, point_alpha = 0.7, alpha = 0.5, size = .25
  ) +
  scale_fill_manual(
    name = "Probability", values = alpha(c("#333333", "#CCCCCC", "#CCCCCC", "#333333"), 0.8),
    labels = c("(0, 0.025]", "(0.025, 0.5]", "(0.5, 0.975]", "(0.975, 1]")
  ) +
  labs(x = "Day of Year", y = "Smolt Migration Year") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2)
  ) +
  scale_x_continuous(breaks = DOYz_xaxis, labels = DOY_lab)
```


```{r p_continuouscovariates_DOY,fig.height= 4, fig.width=6,dev= c('png', 'pdf'),fig.cap="Dashed line represents median value: 122 (DOY)",  cache=TRUE}
ggplot(DATA, aes(x = as.factor(SYage1), y = SYage1_DOY), group = as.factor(SYage1)) +
  geom_point(shape = 21, fill = "transparent", color = "#ABABAB", stroke = .1, size = 1) +
  geom_boxplot(
    fill = "transparent", outlier.fill = "transparent",
    outlier.color = "#ABABAB", outlier.stroke = .1,
    outlier.size = 1, outlier.shape = 21
  ) +
  labs(
    y = "DOY",
    x = "Smolt migration year",
    fill = NULL
  ) +
  geom_hline(yintercept = median(DATA$SYage1_DOY), color = "black", linetype = "dashed") +
  scale_y_continuous(breaks = seq(80, 160, by = 20), limits = c(80, 160)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
```

# Freshwater Covariates {-}
## Figure S7a. Temp
```{r p_continuouscovariates_temp_a,fig.height= 4, fig.width=6,dev= c('png', 'pdf'),fig.cap=" Dashed line represents median value: 9.95 ",  cache=TRUE}
# pdf(here("results", "figures", "manuscript", "FigS7a temperature.pdf"), width=7, height=9, onefile=TRUE)
# LGR.data.temp <- LGR.data[, -3] 
# LGR_temp_long <- data.frame(LGR.data.temp)
# LGR_temp_long <- LGR_temp_long[LGR_temp_long$CY %in% 1998:2020, ]
# LGR_temp_long <- LGR_temp_long[LGR_temp_long$DOY %in% 80:160, ]
# LGR_temp_long_noY <- data.frame(LGR.data.temp[, -1])
# 
# 
# 
# ggplot(LGR_temp_long, aes(x = DOY, y = LGR.temp.1)) +
#   xlim(80, 160) +
#   ylim(2, 18) +
#   geom_point(data = LGR_temp_long_noY, colour = "#DEDEDE", size = 0.25) +
#   geom_point(size = 0.3, colour = pal_cov[2]) +
#   geom_line(colour = pal_cov[2]) +
#   facet_wrap(~CY) +
#   labs(x = "Day of year", y = expression(paste("River temperature (", degree, "C)"))) +
#   theme_bw() +
#   theme(
#     axis.line.x = element_line(linewidth = 0.2),
#     axis.ticks = element_line(linewidth = 0.2),
#     axis.line.y = element_line(linewidth = 0.2)
#   )


#LGR.data seems to be pulling from 1_covariate_data.RMD prior to data join with smolt migration year --switched to using final data set that already has filters above. See plot below. 
ggplot(DATA, aes(x = SYage1_DOY, y = LGR.temp.7d, group=as.factor(SYage1))) +
  xlim(80, 160) +
   ylim(2, 18) +
  geom_point(size = 0.25, color = pal_cov[2] ) +
  geom_line(colour = pal_cov[2]) +
  gghighlight::gghighlight( )+
  labs(x = "Day of year", y = expression(paste("River temperature (", degree, "C)"))) +
  facet_wrap(~SYage1) + 
  theme_bw() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2)
  )
```


```{r p_continuouscovariates_temp,fig.height= 4, fig.width=6,dev= c('png', 'pdf'),fig.cap=" Dashed line represents median value: 9.95 ",  cache=TRUE}
# boxplot
# pdf(here("results", "figures", "manuscript", "FigS7b temperature.pdf"), width=7, height=5, onefile=TRUE)

ggplot(DATA, aes(x = as.factor(SYage1), y = LGR.temp.7d), group = as.factor(SYage1)) +
  geom_point(shape = 21, fill = "transparent", color = "#ABABAB", stroke = .1, size = 1) +
  geom_boxplot(
    fill = "transparent", outlier.fill = "transparent",
    outlier.color = "#ABABAB", outlier.stroke = .1,
    outlier.size = 1, outlier.shape = 21
  ) +
  labs(
    y = expression("Temperature"),
    x = "Smolt migration year",
    fill = NULL
  ) +
  geom_hline(yintercept = median(DATA$LGR.temp.1), linetype = "dashed") +
  scale_y_continuous(breaks = seq(5, 15, 5), limits = c(5, 15), labels = ~ paste0(.x, "°C")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
```



## Figure S8a. Flow

```{r p_continuouscovariates_flow_a,fig.height= 4, fig.width=6,dev= c('png', 'pdf'),fig.cap="Dashed line represents median value: 89.51 kcfs", cache=TRUE}
# pdf(here("results", "figures", "manuscript", "FigS8a flow.pdf"), width=7, height=9, onefile=TRUE)
# LGR.data.flow <- LGR.data[, -4]
# LGR_flow_long <- data.frame(LGR.data.flow)
# LGR_flow_long <- LGR_flow_long[LGR_flow_long$CY %in% 1998:2020, ]
# LGR_flow_long <- LGR_flow_long[LGR_flow_long$DOY %in% 80:160, ]
# LGR_flow_long_noY <- data.frame(LGR.data[, -1])
# 
# LGR_flow_long$CY<-as.numeric(LGR_flow_long$CY)
# LGR_flow_long$DOY<-as.numeric(LGR_flow_long$DOY)
# LGR_flow_long_noY$DOY<-as.numeric(LGR_flow_long_noY$DOY)
#   mutate(DOY = as.numeric(DOY))
# 
# ggplot(LGR_flow_long, aes(x = DOY, y = LGR.flow.1)) +
#   xlim(80, 160) +
#   geom_point(data = LGR_flow_long_noY, colour = "#DEDEDE", size = 0.25) +
#   geom_point(size = 0.3, colour = pal_cov[2]) +
#   geom_line(colour = pal_cov[2]) +
#   facet_wrap(~CY) +
#   labs(x = "Day of year", y = "River flow (kcfs)") +
#   theme_bw() +
#   theme(
#     axis.line.x = element_line(linewidth = 0.2),
#     axis.ticks = element_line(linewidth = 0.2),
#     axis.line.y = element_line(linewidth = 0.2)
#   )

###same as temp, updated to using LGR.flow used in our dataset
ggplot(DATA, aes(x = SYage1_DOY, y = LGR.flow.7d, group=as.factor(SYage1))) +
  xlim(80, 160) +
  geom_point(size = 0.25, color = pal_cov[2] ) +
  geom_line(colour = pal_cov[2]) +
  gghighlight::gghighlight( )+
  labs(x = "Day of year", y = "River flow (kcfs)") +
  facet_wrap(~SYage1) + 
  theme_bw() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2)
  )

```


```{r p_continuouscovariates_flow,fig.height= 4, fig.width=6,dev= c('png', 'pdf'),fig.cap="Dashed line represents median value: 89.51 kcfs",  cache=TRUE}
# boxplot
# pdf(here("results", "figures", "manuscript", "FigS8b flow.pdf"), width=7, height=5, onefile=TRUE)
ggplot(DATA, aes(x = as.factor(SYage1), y = LGR.flow.7d), group = as.factor(SYage1)) +
  geom_point(shape = 21, fill = "transparent", color = "#ABABAB", stroke = .1, size = 1) +
  geom_boxplot(
    fill = "transparent", outlier.fill = "transparent",
    outlier.color = "#ABABAB", outlier.stroke = .1,
    outlier.size = 1, outlier.shape = 21
  ) +
  labs(
    y = "Flow (kcfs)",
    x = "Smolt migration year",
    fill = NULL
  ) +
  geom_hline(yintercept = median(DATA$LGR.flow.7d),  linetype = "dashed") +
  scale_y_continuous(breaks = seq(25, 200, 25), limits = c(25, 200)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
```


# Figure S9. Marine indices 

```{r cov_dataimport, include=FALSE}
#### NPGO raw data
NPGO.raw.data <- fread("https://www.o3d.org/npgo/data/NPGO.txt",
  skip = 30,
  header = FALSE,
  fill=TRUE
)
NPGO.raw.data <- NPGO.raw.data[1:889,]

NPGO.ONDJFM.T <- NPGO.raw.data %>%
  rename(year = 1, month = 2, NPGO = 3) %>%
  mutate(CY = as.numeric(year)) %>%
  mutate(BYage0 = ifelse(month %in% c(1:9), CY - 1, CY)) %>%
  mutate(
    SYage1 = BYage0,
    AYageT = BYage0 + 1
  ) %>%
#  mutate(across(!NPGO, as.character)) %>% 
 filter(month %in% c(10:12, 1:3),
        between (SYage1, 1998, 2023)) %>%
  select(SYage1, month, NPGO)


# 3-year NPGO
NPGO.raw.data <- NPGO.raw.data %>%
  rename(year = 1, month = 2, NPGO = 3) %>%
  mutate(CY = as.numeric(year)) %>%
  mutate(
    BYage0 = ifelse(month %in% c(1:9), CY - 1, CY),
    WY = ifelse(month %in% c(1:9), CY, (CY + 1))
  ) %>%
  mutate(
    SYage1 = BYage0,
    AYageT = BYage0 +1
  ) %>%
  mutate(across(!NPGO, as.character))

NPGO.ONDJFM.T.3 <- NPGO.raw.data %>%
  filter(month %in% c(10:12, 1:3)) %>%
  summarise(.by = SYage1, NPGO.ONDJFM.T = round(mean(NPGO), 3))
NPGO.ONDJFM.T.3$NPGO.ONDJFM.T.AVG3 <- rollmean(NPGO.ONDJFM.T.3$NPGO.ONDJFM.T, 
                                               k=3, align="left", fill=c(NA,NA,NA))
NPGO.ONDJFM.T.3 <- NPGO.ONDJFM.T.3[,-2]



### SSTarc raw data
SSTarc.raw.data <- read_csv(here("data", "covariate data", "ersstArc.anom.csv")) # data is not public

sstarc.ONDJFM.T <- SSTarc.raw.data %>%
  mutate(CY = as.numeric(year)) %>%
  mutate(
    BYage0 = ifelse(month %in% c(1:9), CY - 1, CY),
    WY = ifelse(month %in% c(1:9), CY, (CY + 1))
  ) %>%
  mutate(
    SYage1 = BYage0,
    AYageT = BYage0 + 1
  ) %>%
#  mutate(across(!sstarc, as.character)) %>% 
  filter(month %in% c(10:12, 1:3),
         between (AYageT, 1998, 2023)) %>%
  select(AYageT, month, sstarc)

# # join data
# df.ocean <- inner_join(NPGO.ONDJFM.T, sstarc.ONDJFM.T, by = c("AYageT", "month") ) %>% 
#   filter(between(AYageT, 1998, 2023))

```


```{r p_NPGO_SSTarc,fig.height=4, fig.width=7,dev= c('png', 'pdf')}
# 
# dum.NPGO.ONDJFM.T <- NPGO.ONDJFM.T[NPGO.ONDJFM.T$AYageT %in% 1998:2023, ]
# NPGO.ONDJFM.T.month <- data.frame(dum.NPGO.ONDJFM.T,
#   month = c(10:12, 1:3)
# )

# NPGO
p_raw_npgo <- ggplot(NPGO.ONDJFM.T, aes(x = SYage1, y = NPGO)) +
  geom_point(shape = rep(c("O", "N", "D", "J", "F", "M"), 26), size = 2) +
  stat_summary(
    geom = "line", group = 1, fun = "mean",
    size = 1, colour = pal_cov[3], alpha = 0.5
  ) +
  stat_summary(
    geom = "point", group = 1, fun = "mean",
    shape = 19, size = 3, col = pal_cov[3], alpha = 0.7
  ) +
  scale_x_continuous(breaks = seq(1998,2023, 1))+
  labs(x = "", y = "NPGO index") +
  ggtitle("(a)") +
  theme_minimal() +
  theme(
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2)
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1),
    plot.title = element_text(size = 10)
  )

# 3-year NPGO
NPGO.ONDJFM.T.3.ggplot <- NPGO.ONDJFM.T.3[NPGO.ONDJFM.T.3$SYage1 %in% 1998:2023,]
NPGO.ONDJFM.T.3.ggplot[,1] <- as.numeric(NPGO.ONDJFM.T.3.ggplot[,1])

p_npgo3 <- ggplot(NPGO.ONDJFM.T.3.ggplot, aes(x = SYage1, y = NPGO.ONDJFM.T.AVG3)) +
  geom_line(size = 1, colour = pal_cov[3], alpha = 0.5) +
  geom_point(shape = 19, size = 3, col = pal_cov[3], alpha = 0.7) +
  scale_x_continuous(breaks = seq(1998,2023, 1))+
  labs(x = expression("Year" ~ italic(T) ), 
       y = "3-year mean of NPGO index") +
  ggtitle("(b)") +
  theme_minimal() +
  theme(
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2)
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1),
    axis.title.x = element_text(margin = margin(t = 8, r = 0, b = 0, l = 0)),
    plot.title = element_text(size = 10)
  )


# SSTarc
# dum.sstarc.data <- sstarc.data[sstarc.data$year %in% 1997:2023, ]
# dum.sstarc.data <- dum.sstarc.data[dum.sstarc.data$month %in% c(10:12, 1:3), ]
# sstarc.data.T.month <- dum.sstarc.data[-(1:3), c("AYageT", "month", "sstarc")]

p_raw_SSTarc <- ggplot(sstarc.ONDJFM.T, aes(x = AYageT, y = sstarc)) +
  geom_point(shape = rep(c("O", "N", "D", "J", "F", "M"), 26), size = 2) +
  stat_summary(
    geom = "line", group = 1, fun = "mean",
    size = 1, colour = pal_cov[3], alpha = 0.5
  ) +
  stat_summary(
    geom = "point", group = 1, fun = "mean",
    shape = 19, size = 3, col = pal_cov[3], alpha = 0.7
  ) +
  scale_x_continuous(breaks = seq(1998,2023, 1))+
  labs(
    x = expression("Year" ~ italic(T) ~ "- 1"),
    y = expression(paste(SST[ARC], " index"))
  ) +
  ggtitle("(b)") +
  theme_minimal() +
  theme(
    axis.line.x = element_line(linewidth = 0.2),
    axis.ticks = element_line(linewidth = 0.2),
    axis.line.y = element_line(linewidth = 0.2)
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1),
    plot.title = element_text(size = 10)
  )


# pdf(here("results", "figures", "manuscript", "FigS10 NPGO.pdf"), width=7, height=8, onefile=TRUE)

patch3 <- p_raw_npgo / p_npgo3
patch3
```
